---
title: "Homework 2"
author: Yujie Li
date: 2024-09-30 
output: 
  html_document:
    toc: true 
    toc_float: true 
   
output: github_document 
---


```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(janitor)
library(readr)
```


## Problem 1 
```{r}
nyc_ts= 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", show_col_types = FALSE) %>%
  janitor::clean_names() |> 
  select(line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, 
         entry, vending, entrance_type, ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

nyc_ts
```

## This dataset contains varibales as followed: line, station_name, station_latitude, station_longitude, route1, route2,  route3, route4, route5, entry, vending, entrance_type, and ada. 
## I used "select" to clean the data by retaining the relevant columns and "mutate" to convert the 'entry' variable with "YES" and "NO" into a logical format (TRUE/FALSE). 
## The dimension of the resulting dataset is 1868 x 13 (rows x columns). 
## These data are relatively tidy comparing to the original one, but the routes served could be further tidied (there are too many columns for routes).















# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.


```{r learning_assessment_1}
library(tidyverse)

la_df = tibble(
  norm_samp = rnorm(500, mean = 1),
  norm_samp_pos = norm_samp > 0,
  abs_norm_samp = abs(norm_samp)
)

ggplot(la_df, aes(x = abs_norm_samp)) + geom_histogram()

median_samp = median(pull(la_df, norm_samp))
```

The median of the variable containing absolute values is 
`r round(median_samp, digits = 2)`.

```{r}
la_df = 
  tibble(
    norm_samp = rnorm(n = 500, mean = 1),
    samp_g0 = norm_samp > 0,
    obs_v_samp = obs(norm_samp)
  )

  ggplot(la_df, aes(x = abs_norm_samp)) + 
    geom_histogram()
```

text formatting 
------------------------------------------------------------

*italic*  or  _italic_
**bold**  or  __bold__
`code`
superscript^2^ and subscript~2~


Headings
------------------------------------------------------------

# 1st Level Header

## 2nd Level Header

### 3rd Level Header


Lists
------------------------------------------------------------

*   Bulleted list item 1

*   Item 2

    * Item 2a

    * Item 2b

1.  Numbered list item 1

1.  Item 2. The numbers are incremented automatically in the output.


Tables 
------------------------------------------------------------

First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell